```{r, echo=FALSE, hide=TRUE}
library('tidyverse')
```

# Project Descriotion
My project aims to create a statistical analysis of the spotify data from tidytuesday, which contains various features 
of songs such as danceability, energy, tempo, and loudness.
These features provide valuable insights into the audio characteristics of popular music. Using this dataset, we aim to explore 
relationships between different audio features and uncover patterns in song popularity, 
genre trends, and listener preferences. The final output will include a "killer graph" that highlights key findings in a visually 
compelling way, demonstrating how specific audio features contribute to the popularity of songs on the Spotify platform.
The main object is finding key elements among different songs to find what contributes to the popularity of a song.

# What to find with the "killer graph"
The "killer graph" for this project aims to reveal how song characteristics correlate with their popularity. 
Specifically, the goal is to visualize clusters or trends that show which combination of audio 
features is most likely to produce a hit song.
For example, a heatmap or scatterplot showing how danceability and energy impact song popularity might offer intriguing 
insights into modern music trends.

# Retrieving the data

```{r}
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
head(spotify_songs)
```

# Verify dimension and variables

```{r}
variable_num <- length(names(spotify_songs))
print(paste("There are total of ",variable_num, "variables"))
```

All the variable names:

```{r}
names(spotify_songs)
```


```{r}
n <- nrow(spotify_songs)
print(paste("There are total of ",n,"observations in the dataset"))
```



# Saving the dataset

```{r}
save(spotify_songs, 
     file = "RData/spotify.RData ")
```





